---
title: "Functionalizing Rasters from NEON"
author: "Megan Cattau"
date: "June 21, 2016"
output: html_document
---

## Objectives

Set the working directory and load packages
```{r set-wd-load-packages}
setwd("~/Documents/data/NEONDI-2016/Tues_pm_functional")
library(raster)
# library(rhdf5)
# library(rgdal)
```

Import a raster — A lidar canopy height model (CHM) and set values == 0 to NA (not trees)

```{r import-CHM-and-set-0-to-NA}
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

# assign chm values of 0 to NA
chm[chm==0] <- NA
```

Visualize histogram/density and plot vertical cutoff lines.

```{r hist-and-cutoffs}

#inspect data
summary(chm)
# select thresholds 9 (1st qu) and 16 (Median)
density(chm)
hist(chm)
abline(v=c(9,16), col="red")

```

Classify the raster: <9m is generally grasses / understory, <16m small trees, and the rest are tall trees. A function could import the desired thresholds. 


```{r classify}
class.m <- c(0, 9, 1,
             9, 16, 2,
             16, 60, 3)
class.m

# reshape the object into a matrix with columns and rows
rcl.m <- matrix(class.m, 
                ncol=3, 
                byrow=TRUE)
rcl.m

# reclassify the raster chm using the reclass function to create a new raster.
chm.classify <- reclassify(chm, 
                     rcl.m)

# look hist w function
hist(chm)
abline(v=rcl.m[,2], col="red")
```

Plot it

```{r plot}
# make room for a legend
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
# plot
plot(chm.classify,
     col=c("green","blue", "red"),
     main="Classified CHM",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Grasses / understory", "Short trees", "Tall trees"),  # make sure the order matches the colors, next
       fill = c("green", "blue", "red"),
       bty="n") # turn off border
par(xpd=FALSE)
```

Export Classified Raster
```{r export, eval=FALSE}
# export geotiff
writeRaster(chm.classify,
            filename="../outputs/TEAK/Classified_CHM.tif",
            format="GTiff",
            options="COMPRESS=LZW",
            overwrite = TRUE,
            NAflag = -9999)
```

Take the chm and create a hillshade (http://neon-workwithdata.github.io/neon-data-institute-2016/R/create-hillshade-R/)
PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
Export the plot figure to a pdf – publishable
Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.